---
title: "Testing Posterior Predictive Model Checking for the Time-Varying Dynamic Partial Credit Model - Part 2"
author: "Sebastian Castro-Alvarez"
date: "`r format(Sys.Date(), '%B %d del %Y')`"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{amsmath}
output: 
  bookdown::pdf_document2:
    toc: false
    keep_tex: false
    number_section: false
params:
  seed: 1234
  nitems: 6
bibliography: references.bib
csl: apa7.csl
link-citations: true
always_allow_html: true
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE, fig.height = 4, fig.width = 6, fig.pos = "!H",  
                      warning=FALSE, message=FALSE)

# seeds  <- 1002:1005
# Nitems <- c(3)
# for (i in 1:length(seeds)) {
#   for (j in 1:length(Nitems)) {
#     seed   <- seeds[i]
#     nitems <- Nitems[j]
#       rmarkdown::render("Rmarkdown/PPMC_Tests_Part2.Rmd", 
#                         params = list(seed = seed, nitems = nitems), 
#                         output_file = paste0("PPMC_Tests_Part2_", seed, "_", nitems))
#   }
# }
```

```{r pre-env}
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores() * (1/2))
library(bayesplot)
color_scheme_set("darkgray")

source("../R/IRT_models.R")
source("../R/IRT_plots.R")
source("../R/PPMC.R")
source("../R/genTVDPCM.R")
source("../R/tvdpcm2stan.R")

seed <- params$seed
set.seed(seed)

model <- stan_model(file = "../Stan/tv_dpcm_int_v5.1.stan", verbose = FALSE)
```

In this document, we generate data that should show model misfit after fitting the TV-DPCM, which integrates the partial credit model [@Masters2016] and the time-varying autoregressive model [@Bringmann2017]. In what follows, we describe the generating data sets and try to relate the responses to real life situations, then we fit the TV-DPCM, and then we compute several PPMC methods propose for IRT models and some modified versions of these methods.

```{r fixvalues}
nT     <- 200           # Number of time points
I      <- params$nitems # Number of items
K      <- 5             # Number of categories per item
M      <- K - 1         # Number of thresholds per item
lambda <- 0.5           # Size of the autoregressive effect
in_var <- 1             # Variance of the innovations
```

\clearpage

## Change in Meaning

In this first example, we generate responses of one subject for 200 time points and `r I` items measured on a 5-point Likert scale. However, the responses to one item are reversed for the last third of the measurements. This aims to reflect situations in which the meaning of an item changes over time. In particular, consider the following situation, a subject filled in a ESM questionnaire while they were following psychotherapy to treat depression. We have their responses to the items of positive affect, one of which was "I am feeling energetic". At the start, feeling energetic was associated with feeling happier. However, after psychotherapy, the subject's depressive symptoms actually improve and for the last third, not feeling energetic is actually associated with positive emotions. So, now when the person is feeling happy, their responses on feeling energetic are actually low. To represent this, we reverse the simulated scores of item `r I` of the last third of the responses. The time series of the observed sumscores are presented in Figure \@ref(fig:ts1).

```{r}
# Generate Data based on the TV-DPCM
dataTRUE <- gen.TVDPCM(nT = nT, 
                       I  = I, 
                       K  = K, 
                       pop.param = list(lambda = lambda), 
                       seed = seed, 
                       FUN  = "logarithmic",
                       maxAbsValue = 0.5)

responses <- dataTRUE$data
responses[ceiling(2*nT/3):nT, I] <- 6 - responses[ceiling(2*nT/3):nT, I] 
```

```{r ts1, fig.cap = "Time Series of the Sumscores"}
plot(rowSums(responses), type = "l", ylab = "sumscores", xlab = "Time",
     las = 1, lwd = 2)
```

```{r}
standata <- tvdpcm2stan_data(resp = as.matrix(responses),
                             I    = I,
                             K    = K,
                             nT   = nT,
                             n_knots  = 8,
                             s_degree = 3)
```

```{r}
tvdpcm_inits <- function() {
              list(lambda = runif(1, -1, 1),
                   beta   = array(rnorm(I * (K - 1), 0, 3), dim = c(I, K - 1)),
                   inno   = rnorm(nT, 0, 3),
                   sigma  = rlnorm(1, 1))
            }

begin.time <- proc.time()
fit <- sampling(model,                  # Stan model. 
                data = standata,        # Data.
                iter = 2000,            # Number of iterations.
                chains  = 3,            # Number of chains.
                warmup  = 500,          # Burn-in samples.
                init    = tvdpcm_inits, # Initial values
                seed = seed,            # Seed
                pars = c("beta", "theta", "lambda",
                         "sigma2", "pvar", "attractor", "rep_y"),
                control = list(adapt_delta   = 0.99,
                               max_treedepth = 15) # Other parameters to control sampling behavior.
                  ) 
run.time <- proc.time() - begin.time
rm(begin.time)
```

```{r}
stan.diag <- monitor(extract(fit, 
                             pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"),
                             permuted = FALSE, inc_warmup = FALSE), 
                                 warmup = 0, print = FALSE)
            
ndiv  <- get_num_divergent(fit)     # number of divergent transitions
nbfmi <- get_low_bfmi_chains(fit)   # number of chains with a low bayesian fraction missing information 
ntree <- get_num_max_treedepth(fit) # number of transitions that exceeded the maximum treedepth
nbulk <- sum(stan.diag$Bulk_ESS < 100 * dim(fit)[2]) # number of parameters with low bulk ESS
ntail <- sum(stan.diag$Tail_ESS < 100 * dim(fit)[2]) # number of parameters with low tail ESS
          
            #max Rhat
maxRhat <- round(max(rhat(fit, pars = c("beta", "theta","lambda",
                                      "sigma2", "pvar", "attractor"))), 4)
nRhat   <- sum(rhat(fit, pars = c("beta", "theta", "lambda",
                                  "sigma2", "pvar", "attractor")) > 1.05)
```

The HMC diagnostics of the fitting model are presented in Table \@ref(tab:diagtable1). The model worked well if all the values in this table are 0.

```{r diagtable1}
diag.table <- matrix(c(nRhat, ndiv, length(nbfmi), ntree, nbulk, ntail), nrow = 1)
row.names(diag.table)  <- "Fitted Model"
colnames(diag.table) <- c("Rhat>1.05", "N. Divergent", "N. Low BFMI", 
                           "Excedeed Treedepth", "Low Bulk ESS", "Low Tail ESS")
kbl(diag.table, align = "c", booktabs = TRUE, caption = "HMC Diagnostics", escape = FALSE)
```

\clearpage

### Posterior Predictive Model Checking Methods

Next, the following tables present the computed posterior predictive p-values based on multiple PPMC methods.

```{r, results = "hide", fig.show = "hide"}
ppmc01 <- ppmc.acf(object = fit, data = standata, lag.max = 3)
ppmc02 <- ppmc.racf(object = fit, data = standata)
ppmc03 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE, sumscores = TRUE)
ppmc04 <- ppmc.mssd(object = fit, data = standata)
ppmc05 <- ppmc.itcor(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc06 <- ppmc.itcor2(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc07 <- ppmc.itcor3(object = fit, data = standata, quiet = TRUE)
ppmc08 <- ppmc.Q1(object = fit, data = standata, quiet = TRUE)
ppmc09 <- ppmc.Q1.alt(object = fit, data = standata, quiet = TRUE)
ppmc10 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE)
ppmc11 <- ppmc.Q3(object = fit, data = standata)$ppp
ppmc12 <- ppmc.OR(object = fit, data = standata)$ppp
ppmc13 <- ppmc.ORDiff(object = fit, data = standata)$ppp
ppmc14 <- ppmc.cov.resid(object = fit, data = standata)$ppp
ppmc15 <- ppmc.cov.rediff(object = fit, data = standata)$ppp
```

```{r}
ppmctest <- matrix(c(ppmc01, ppmc02, ppmc03, ppmc04, rep(NA, 6)), ncol = 2)
ppmctest[, 2] <- as.numeric(ppmctest[, 1] <= 0.05 | ppmctest[, 1] >= 0.95)
row.names(ppmctest)  <- c("ACF-1", "ACF-2", "ACF-3", "PACF", "LPACF", "MSSD")
colnames(ppmctest) <- c("ppp", "Misfit")
kbl(ppmctest, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Test Level", escape = FALSE)
```

```{r}
ppmcitem <- matrix(c(ppmc05, ppmc06, ppmc07, 
                     ppmc08, ppmc09, ppmc10), ncol = I, byrow = TRUE)
misfit   <- apply(ppmcitem, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcitem <- cbind(ppmcitem, misfit)
row.names(ppmcitem)  <- c("IT-cor", "IT-cor2", "IT-cor3", 
                          "Yens Q1", "Yens Q1 Alt", "ILPACF")
colnames(ppmcitem) <- c(paste0("Item", 1:I), "Misfit")
kbl(ppmcitem, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item Level", escape = FALSE)
```

```{r}
ppmcpairs <- matrix(c(ppmc11, ppmc12, ppmc13, 
                      ppmc14, ppmc15), ncol = (I * (I - 1))/2, byrow = TRUE)
misfit   <- apply(ppmcpairs, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcpairs <- cbind(ppmcpairs, misfit)
row.names(ppmcpairs)  <- c("Yens Q3", "OR", "OR-Diff", 
                          "RESID", "REDIFF")
pairs.index <- which(lower.tri(diag(I)), arr.ind = TRUE)
pairs.index <- apply(pairs.index, 1, function(x) paste(rev(x), collapse = ","))
colnames(ppmcpairs) <- c(paste0("Items(", pairs.index, ")"), "Misfit")
kbl(t(ppmcpairs), align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item-Pairs Level", escape = FALSE)
```

```{r}
out_table <- data.frame(seed, 
                        model = "Meaning",
                        diag.table, 
                        ACF = t(ppmc01),
                        PACF = ppmc02,
                        LPACF = ppmc03,
                        MSSD = ppmc04,
                        itcor = t(ppmc05),
                        itcor2 = t(ppmc06),
                        itcor3 = t(ppmc07),
                        q1 = t(ppmc08),
                        q1alt = t(ppmc09),
                        lpacf = t(ppmc10),
                        t(ppmc11),
                        t(ppmc12),
                        t(ppmc13),
                        t(ppmc14),
                        t(ppmc15))
```


```{r}
rm(diag.table, fit, ppmcitem, ppmcpairs, ppmctest, responses, stan.diag,
   standata, maxRhat, misfit, nbfmi, nbulk, ndiv, nRhat, ntail, ntree,
   pairs.index, ppmc01, ppmc02, ppmc03, ppmc04, ppmc05, ppmc06, ppmc07, 
   ppmc08, ppmc09, ppmc10, ppmc11, ppmc12, ppmc13, ppmc14, ppmc15, run.time)
```

\clearpage

## Default Responses

In this second example, we generate responses of one subject assuming that the subject gets bored with the daily measurements and starts filling their responses with the same answers all the time. In this situation, the participant only chose the response category of the middle for the last third of the study. The generated sumscores are presented in Figure \@ref(fig:ts2).

```{r}
# Generate Data based on the TV-DPCM
dataTRUE <- gen.TVDPCM(nT = nT, 
                       I  = I, 
                       K  = K, 
                       pop.param = list(lambda = lambda), 
                       seed = seed, 
                       FUN  = "logarithmic",
                       maxAbsValue = 0.5)

responses <- dataTRUE$data
responses[ceiling(2*nT/3):nT, ] <- 3 
```

```{r ts2, fig.cap = "Time Series of the Sumscores"}
plot(rowSums(responses), type = "l", ylab = "sumscores", xlab = "Time",
     las = 1, lwd = 2)
```

```{r}
standata <- tvdpcm2stan_data(resp = as.matrix(responses),
                             I    = I,
                             K    = K,
                             nT   = nT,
                             n_knots  = 8,
                             s_degree = 3)
```

```{r}
tvdpcm_inits <- function() {
              list(lambda = runif(1, -1, 1),
                   beta   = array(rnorm(I * (K - 1), 0, 3), dim = c(I, K - 1)),
                   inno   = rnorm(nT, 0, 3),
                   sigma  = rlnorm(1, 1))
            }

begin.time <- proc.time()
fit <- sampling(model,                  # Stan model. 
                data = standata,        # Data.
                iter = 2000,            # Number of iterations.
                chains  = 3,            # Number of chains.
                warmup  = 500,          # Burn-in samples.
                init    = tvdpcm_inits, # Initial values
                seed = seed,            # Seed
                pars = c("beta", "theta", "lambda",
                         "sigma2", "pvar", "attractor", "rep_y"),
                control = list(adapt_delta   = 0.99,
                               max_treedepth = 15) # Other parameters to control sampling behavior.
                  ) 
run.time <- proc.time() - begin.time
rm(begin.time)
```

```{r}
stan.diag <- monitor(extract(fit, 
                             pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"),
                             permuted = FALSE, inc_warmup = FALSE), 
                                 warmup = 0, print = FALSE)
            
ndiv  <- get_num_divergent(fit)     # number of divergent transitions
nbfmi <- get_low_bfmi_chains(fit)   # number of chains with a low bayesian fraction missing information 
ntree <- get_num_max_treedepth(fit) # number of transitions that exceeded the maximum treedepth
nbulk <- sum(stan.diag$Bulk_ESS < 100 * dim(fit)[2]) # number of parameters with low bulk ESS
ntail <- sum(stan.diag$Tail_ESS < 100 * dim(fit)[2]) # number of parameters with low tail ESS
          
            #max Rhat
maxRhat <- round(max(rhat(fit, pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"))), 4)
nRhat   <- sum(rhat(fit, pars = c("beta", "theta", "lambda",
                                  "sigma2", "pvar", "attractor")) > 1.05)
```

The HMC diagnostics of the fitting model are presented in Table \@ref(tab:diagtable2). The model worked well if all the values in this table are 0.

```{r diagtable2}
diag.table <- matrix(c(nRhat, ndiv, length(nbfmi), ntree, nbulk, ntail), nrow = 1)
row.names(diag.table)  <- "Fitted Model"
colnames(diag.table) <- c("Rhat>1.05", "N. Divergent", "N. Low BFMI", 
                           "Excedeed Treedepth", "Low Bulk ESS", "Low Tail ESS")
kbl(diag.table, align = "c", booktabs = TRUE, caption = "HMC Diagnostics", escape = FALSE)
```

\clearpage

### Posterior Predictive Model Checking Methods

Next, the following tables present the computed posterior predictive p-values based on multiple PPMC methods.

```{r, results = "hide", fig.show = "hide"}
ppmc01 <- ppmc.acf(object = fit, data = standata, lag.max = 3)
ppmc02 <- ppmc.racf(object = fit, data = standata)
ppmc03 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE, sumscores = TRUE)
ppmc04 <- ppmc.mssd(object = fit, data = standata)
ppmc05 <- ppmc.itcor(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc06 <- ppmc.itcor2(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc07 <- ppmc.itcor3(object = fit, data = standata, quiet = TRUE)
ppmc08 <- ppmc.Q1(object = fit, data = standata, quiet = TRUE)
ppmc09 <- ppmc.Q1.alt(object = fit, data = standata, quiet = TRUE)
ppmc10 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE)
ppmc11 <- ppmc.Q3(object = fit, data = standata)$ppp
ppmc12 <- ppmc.OR(object = fit, data = standata)$ppp
ppmc13 <- ppmc.ORDiff(object = fit, data = standata)$ppp
ppmc14 <- ppmc.cov.resid(object = fit, data = standata)$ppp
ppmc15 <- ppmc.cov.rediff(object = fit, data = standata)$ppp
```

```{r}
ppmctest <- matrix(c(ppmc01, ppmc02, ppmc03, ppmc04, rep(NA, 6)), ncol = 2)
ppmctest[, 2] <- as.numeric(ppmctest[, 1] <= 0.05 | ppmctest[, 1] >= 0.95)
row.names(ppmctest)  <- c("ACF-1", "ACF-2", "ACF-3", "PACF", "LPACF", "MSSD")
colnames(ppmctest) <- c("ppp", "Misfit")
kbl(ppmctest, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Test Level", escape = FALSE)
```

```{r}
ppmcitem <- matrix(c(ppmc05, ppmc06, ppmc07, 
                     ppmc08, ppmc09, ppmc10), ncol = I, byrow = TRUE)
misfit   <- apply(ppmcitem, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcitem <- cbind(ppmcitem, misfit)
row.names(ppmcitem)  <- c("IT-cor", "IT-cor2", "IT-cor3", 
                          "Yens Q1", "Yens Q1 Alt", "ILPACF")
colnames(ppmcitem) <- c(paste0("Item", 1:I), "Misfit")
kbl(ppmcitem, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item Level", escape = FALSE)
```

```{r}
ppmcpairs <- matrix(c(ppmc11, ppmc12, ppmc13, 
                      ppmc14, ppmc15), ncol = (I * (I - 1))/2, byrow = TRUE)
misfit   <- apply(ppmcpairs, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcpairs <- cbind(ppmcpairs, misfit)
row.names(ppmcpairs)  <- c("Yens Q3", "OR", "OR-Diff", 
                          "RESID", "REDIFF")
pairs.index <- which(lower.tri(diag(I)), arr.ind = TRUE)
pairs.index <- apply(pairs.index, 1, function(x) paste(rev(x), collapse = ","))
colnames(ppmcpairs) <- c(paste0("Items(", pairs.index, ")"), "Misfit")
kbl(t(ppmcpairs), align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item-Pairs Level", escape = FALSE)
```

```{r}
out_table <- rbind(out_table,
                   data.frame(seed, 
                        model = "Default",
                        diag.table, 
                        ACF = t(ppmc01),
                        PACF = ppmc02,
                        LPACF = ppmc03,
                        MSSD = ppmc04,
                        itcor = t(ppmc05),
                        itcor2 = t(ppmc06),
                        itcor3 = t(ppmc07),
                        q1 = t(ppmc08),
                        q1alt = t(ppmc09),
                        lpacf = t(ppmc10),
                        t(ppmc11),
                        t(ppmc12),
                        t(ppmc13),
                        t(ppmc14),
                        t(ppmc15)))
```

```{r}
rm(diag.table, fit, ppmcitem, ppmcpairs, ppmctest, responses, stan.diag,
   standata, maxRhat, misfit, nbfmi, nbulk, ndiv, nRhat, ntail, ntree,
   pairs.index, ppmc01, ppmc02, ppmc03, ppmc04, ppmc05, ppmc06, ppmc07, 
   ppmc08, ppmc09, ppmc10, ppmc11, ppmc12, ppmc13, ppmc14, ppmc15, run.time)
```

\clearpage

## Random Responses

Now, we assume that the participant got tired of the daily questionnaires but still decided to fill them in randomly. Here, we randomly generate the responses of the last third of the study. The generated sumscores are presented in Figure \@ref(fig:ts3).

```{r}
# Generate Data based on the TV-DPCM
dataTRUE <- gen.TVDPCM(nT = nT, 
                       I  = I, 
                       K  = K, 
                       pop.param = list(lambda = lambda), 
                       seed = seed, 
                       FUN  = "logarithmic",
                       maxAbsValue = 0.5)

responses <- dataTRUE$data
responses[ceiling(2*nT/3):nT, ] <- sample(1:K, length(ceiling(2*nT/3):nT) * I, 
                                          replace = TRUE) 
```

```{r ts3, fig.cap = "Time Series of the Sumscores"}
plot(rowSums(responses), type = "l", ylab = "sumscores", xlab = "Time",
     las = 1, lwd = 2)
```

```{r}
standata <- tvdpcm2stan_data(resp = as.matrix(responses),
                             I    = I,
                             K    = K,
                             nT   = nT,
                             n_knots  = 8,
                             s_degree = 3)
```

```{r}
tvdpcm_inits <- function() {
              list(lambda = runif(1, -1, 1),
                   beta   = array(rnorm(I * (K - 1), 0, 3), dim = c(I, K - 1)),
                   inno   = rnorm(nT, 0, 3),
                   sigma  = rlnorm(1, 1))
            }

begin.time <- proc.time()
fit <- sampling(model,                  # Stan model. 
                data = standata,        # Data.
                iter = 2000,            # Number of iterations.
                chains  = 3,            # Number of chains.
                warmup  = 500,          # Burn-in samples.
                init    = tvdpcm_inits, # Initial values
                seed = seed,            # Seed
                pars = c("beta", "theta", "lambda",
                         "sigma2", "pvar", "attractor", "rep_y"),
                control = list(adapt_delta   = 0.99,
                               max_treedepth = 15) # Other parameters to control sampling behavior.
                  ) 
run.time <- proc.time() - begin.time
rm(begin.time)
```

```{r}
stan.diag <- monitor(extract(fit, 
                             pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"),
                             permuted = FALSE, inc_warmup = FALSE), 
                                 warmup = 0, print = FALSE)
            
ndiv  <- get_num_divergent(fit)     # number of divergent transitions
nbfmi <- get_low_bfmi_chains(fit)   # number of chains with a low bayesian fraction missing information 
ntree <- get_num_max_treedepth(fit) # number of transitions that exceeded the maximum treedepth
nbulk <- sum(stan.diag$Bulk_ESS < 100 * dim(fit)[2]) # number of parameters with low bulk ESS
ntail <- sum(stan.diag$Tail_ESS < 100 * dim(fit)[2]) # number of parameters with low tail ESS
          
            #max Rhat
maxRhat <- round(max(rhat(fit, pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"))), 4)
nRhat   <- sum(rhat(fit, pars = c("beta", "theta", "lambda",
                                  "sigma2", "pvar", "attractor")) > 1.05)
```

The HMC diagnostics of the fitting model are presented in Table \@ref(tab:diagtable3). The model worked well if all the values in this table are 0.

```{r diagtable3}
diag.table <- matrix(c(nRhat, ndiv, length(nbfmi), ntree, nbulk, ntail), nrow = 1)
row.names(diag.table)  <- "Fitted Model"
colnames(diag.table) <- c("Rhat>1.05", "N. Divergent", "N. Low BFMI", 
                           "Excedeed Treedepth", "Low Bulk ESS", "Low Tail ESS")
kbl(diag.table, align = "c", booktabs = TRUE, caption = "HMC Diagnostics", escape = FALSE)
```

\clearpage

### Posterior Predictive Model Checking Methods

Next, the following tables present the computed posterior predictive p-values based on multiple PPMC methods.

```{r, results = "hide", fig.show = "hide"}
ppmc01 <- ppmc.acf(object = fit, data = standata, lag.max = 3)
ppmc02 <- ppmc.racf(object = fit, data = standata)
ppmc03 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE, sumscores = TRUE)
ppmc04 <- ppmc.mssd(object = fit, data = standata)
ppmc05 <- ppmc.itcor(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc06 <- ppmc.itcor2(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc07 <- ppmc.itcor3(object = fit, data = standata, quiet = TRUE)
ppmc08 <- ppmc.Q1(object = fit, data = standata, quiet = TRUE)
ppmc09 <- ppmc.Q1.alt(object = fit, data = standata, quiet = TRUE)
ppmc10 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE)
ppmc11 <- ppmc.Q3(object = fit, data = standata)$ppp
ppmc12 <- ppmc.OR(object = fit, data = standata)$ppp
ppmc13 <- ppmc.ORDiff(object = fit, data = standata)$ppp
ppmc14 <- ppmc.cov.resid(object = fit, data = standata)$ppp
ppmc15 <- ppmc.cov.rediff(object = fit, data = standata)$ppp
```

```{r}
ppmctest <- matrix(c(ppmc01, ppmc02, ppmc03, ppmc04, rep(NA, 6)), ncol = 2)
ppmctest[, 2] <- as.numeric(ppmctest[, 1] <= 0.05 | ppmctest[, 1] >= 0.95)
row.names(ppmctest)  <- c("ACF-1", "ACF-2", "ACF-3", "PACF", "LPACF", "MSSD")
colnames(ppmctest) <- c("ppp", "Misfit")
kbl(ppmctest, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Test Level", escape = FALSE)
```

```{r}
ppmcitem <- matrix(c(ppmc05, ppmc06, ppmc07, 
                     ppmc08, ppmc09, ppmc10), ncol = I, byrow = TRUE)
misfit   <- apply(ppmcitem, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcitem <- cbind(ppmcitem, misfit)
row.names(ppmcitem)  <- c("IT-cor", "IT-cor2", "IT-cor3", 
                          "Yens Q1", "Yens Q1 Alt", "ILPACF")
colnames(ppmcitem) <- c(paste0("Item", 1:I), "Misfit")
kbl(ppmcitem, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item Level", escape = FALSE)
```

```{r}
ppmcpairs <- matrix(c(ppmc11, ppmc12, ppmc13, 
                      ppmc14, ppmc15), ncol = (I * (I - 1))/2, byrow = TRUE)
misfit   <- apply(ppmcpairs, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcpairs <- cbind(ppmcpairs, misfit)
row.names(ppmcpairs)  <- c("Yens Q3", "OR", "OR-Diff", 
                          "RESID", "REDIFF")
pairs.index <- which(lower.tri(diag(I)), arr.ind = TRUE)
pairs.index <- apply(pairs.index, 1, function(x) paste(rev(x), collapse = ","))
colnames(ppmcpairs) <- c(paste0("Items(", pairs.index, ")"), "Misfit")
kbl(t(ppmcpairs), align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item-Pairs Level", escape = FALSE)
```


```{r}
out_table <- rbind(out_table,
                   data.frame(seed, 
                        model = "Random",
                        diag.table, 
                        ACF = t(ppmc01),
                        PACF = ppmc02,
                        LPACF = ppmc03,
                        MSSD = ppmc04,
                        itcor = t(ppmc05),
                        itcor2 = t(ppmc06),
                        itcor3 = t(ppmc07),
                        q1 = t(ppmc08),
                        q1alt = t(ppmc09),
                        lpacf = t(ppmc10),
                        t(ppmc11),
                        t(ppmc12),
                        t(ppmc13),
                        t(ppmc14),
                        t(ppmc15)))
```

```{r}
rm(diag.table, fit, ppmcitem, ppmcpairs, ppmctest, responses, stan.diag,
   standata, maxRhat, misfit, nbfmi, nbulk, ndiv, nRhat, ntail, ntree,
   pairs.index, ppmc01, ppmc02, ppmc03, ppmc04, ppmc05, ppmc06, ppmc07, 
   ppmc08, ppmc09, ppmc10, ppmc11, ppmc12, ppmc13, ppmc14, ppmc15, run.time)
```

\clearpage

## Item Parameter Drift of some Items

In this example, we generate responses where there was item parameter drift for the last third of the study. This means that the item parameters changed. We only modify two of the `r I` items. For one of the items, the parameters are increased and for the other the parameters are decreased. In particular, we increased the parameters of Item `r I - 1` and decreased the parameters of Item `r I`. The generated sumscores are presented in Figure \@ref(fig:ts4).

```{r}
# Generate Data based on the TV-DPCM
dataTRUE <- gen.TVDPCM(nT = nT, 
                       I  = I, 
                       K  = K, 
                       pop.param = list(lambda = lambda), 
                       seed = seed, 
                       FUN  = "logarithmic",
                       maxAbsValue = 0.5)

# Use same theta as in data1
theta <- dataTRUE$theta.gen

# Location
delta <- rowMeans(dataTRUE$thresholds.gen)
  
# Steps
taus  <- dataTRUE$thresholds.gen - delta

# Create item parameters for the last third.
thresholds2 <- dataTRUE$thresholds.gen
thresholds2[I - 1, ] <- thresholds2[I - 1, ] + 2 
thresholds2[I, ] <- thresholds2[I, ] - 2  

# Location
delta2 <- rowMeans(thresholds2)

# Step parameters
taus2 <- thresholds2 - delta2

# Generate responses  half 1
probs.array <- array(NA, dim = c(ceiling(2 * nT/3), I, K))

for (y in 0:M) {
  probs.array[, , y + 1] <- P.GPCM(y     = y, 
                                   alpha = rep(1, I), 
                                   delta = delta, 
                                   taus  = taus, 
                                   theta = theta[1:ceiling(2 * nT/3)], 
                                   M     = M)
}
responses1   <- apply(probs.array, 1:2, function(vec) {which( rmultinom(1, 1, vec) == 1)})
rm(probs.array, y)

# Generate responses  half 2
probs.array <- array(NA, dim = c(nT - ceiling(2 * nT/3), I, K))

for (y in 0:M) {
  probs.array[, , y + 1] <- P.GPCM(y     = y, 
                                   alpha = rep(1, I), 
                                   delta = delta2, 
                                   taus  = taus2, 
                                   theta = theta[(ceiling(2 * nT/3) + 1):nT], 
                                   M     = M)
}
responses2   <- apply(probs.array, 1:2, function(vec) {which( rmultinom(1, 1, vec) == 1)})
rm(probs.array, y)

responses <- rbind(responses1, responses2)
rm(responses1, responses2)
```

```{r ts4, fig.cap = "Time Series of the Sumscores"}
plot(rowSums(responses), type = "l", ylab = "sumscores", xlab = "Time",
     las = 1, lwd = 2)
```

```{r}
standata <- tvdpcm2stan_data(resp = as.matrix(responses),
                             I    = I,
                             K    = K,
                             nT   = nT,
                             n_knots  = 8,
                             s_degree = 3)
```

```{r}
tvdpcm_inits <- function() {
              list(lambda = runif(1, -1, 1),
                   beta   = array(rnorm(I * (K - 1), 0, 3), dim = c(I, K - 1)),
                   inno   = rnorm(nT, 0, 3),
                   sigma  = rlnorm(1, 1))
            }

begin.time <- proc.time()
fit <- sampling(model,                  # Stan model. 
                data = standata,        # Data.
                iter = 2000,            # Number of iterations.
                chains  = 3,            # Number of chains.
                warmup  = 500,          # Burn-in samples.
                init    = tvdpcm_inits, # Initial values
                seed = seed,            # Seed
                pars = c("beta", "theta", "lambda",
                         "sigma2", "pvar", "attractor", "rep_y"),
                control = list(adapt_delta   = 0.99,
                               max_treedepth = 15) # Other parameters to control sampling behavior.
                  ) 
run.time <- proc.time() - begin.time
rm(begin.time)
```

```{r}
stan.diag <- monitor(extract(fit, 
                             pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"),
                             permuted = FALSE, inc_warmup = FALSE), 
                                 warmup = 0, print = FALSE)
            
ndiv  <- get_num_divergent(fit)     # number of divergent transitions
nbfmi <- get_low_bfmi_chains(fit)   # number of chains with a low bayesian fraction missing information 
ntree <- get_num_max_treedepth(fit) # number of transitions that exceeded the maximum treedepth
nbulk <- sum(stan.diag$Bulk_ESS < 100 * dim(fit)[2]) # number of parameters with low bulk ESS
ntail <- sum(stan.diag$Tail_ESS < 100 * dim(fit)[2]) # number of parameters with low tail ESS
          
            #max Rhat
maxRhat <- round(max(rhat(fit, pars = c("beta", "theta", "lambda",
                                      "sigma2", "pvar", "attractor"))), 4)
nRhat   <- sum(rhat(fit, pars = c("beta", "theta", "lambda",
                                  "sigma2", "pvar", "attractor")) > 1.05)
```

The HMC diagnostics of the fitting model are presented in Table \@ref(tab:diagtable4). The model worked well if all the values in this table are 0.

```{r diagtable4}
diag.table <- matrix(c(nRhat, ndiv, length(nbfmi), ntree, nbulk, ntail), nrow = 1)
row.names(diag.table)  <- "Fitted Model"
colnames(diag.table) <- c("Rhat>1.05", "N. Divergent", "N. Low BFMI", 
                           "Excedeed Treedepth", "Low Bulk ESS", "Low Tail ESS")
kbl(diag.table, align = "c", booktabs = TRUE, caption = "HMC Diagnostics", escape = FALSE)
```

\clearpage

### Posterior Predictive Model Checking Methods

Next, the following tables present the computed posterior predictive p-values based on multiple PPMC methods.

```{r, results = "hide", fig.show = "hide"}
ppmc01 <- ppmc.acf(object = fit, data = standata, lag.max = 3)
ppmc02 <- ppmc.racf(object = fit, data = standata)
ppmc03 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE, sumscores = TRUE)
ppmc04 <- ppmc.mssd(object = fit, data = standata)
ppmc05 <- ppmc.itcor(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc06 <- ppmc.itcor2(object = fit, data = standata, quiet = TRUE, method = "pearson")
ppmc07 <- ppmc.itcor3(object = fit, data = standata, quiet = TRUE)
ppmc08 <- ppmc.Q1(object = fit, data = standata, quiet = TRUE)
ppmc09 <- ppmc.Q1.alt(object = fit, data = standata, quiet = TRUE)
ppmc10 <- ppmc.lpacf(object = fit, data = standata, quiet = TRUE)
ppmc11 <- ppmc.Q3(object = fit, data = standata)$ppp
ppmc12 <- ppmc.OR(object = fit, data = standata)$ppp
ppmc13 <- ppmc.ORDiff(object = fit, data = standata)$ppp
ppmc14 <- ppmc.cov.resid(object = fit, data = standata)$ppp
ppmc15 <- ppmc.cov.rediff(object = fit, data = standata)$ppp
```

```{r}
ppmctest <- matrix(c(ppmc01, ppmc02, ppmc03, ppmc04, rep(NA, 6)), ncol = 2)
ppmctest[, 2] <- as.numeric(ppmctest[, 1] <= 0.05 | ppmctest[, 1] >= 0.95)
row.names(ppmctest)  <- c("ACF-1", "ACF-2", "ACF-3", "PACF", "LPACF", "MSSD")
colnames(ppmctest) <- c("ppp", "Misfit")
kbl(ppmctest, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Test Level", escape = FALSE)
```

```{r}
ppmcitem <- matrix(c(ppmc05, ppmc06, ppmc07, 
                     ppmc08, ppmc09, ppmc10), ncol = I, byrow = TRUE)
misfit   <- apply(ppmcitem, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcitem <- cbind(ppmcitem, misfit)
row.names(ppmcitem)  <- c("IT-cor", "IT-cor2", "IT-cor3", 
                          "Yens Q1", "Yens Q1 Alt", "ILPACF")
colnames(ppmcitem) <- c(paste0("Item", 1:I), "Misfit")
kbl(ppmcitem, align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item Level", escape = FALSE)
```

```{r}
ppmcpairs <- matrix(c(ppmc11, ppmc12, ppmc13, 
                      ppmc14, ppmc15), ncol = (I * (I - 1))/2, byrow = TRUE)
misfit   <- apply(ppmcpairs, 1, function(x) sum(x <= 0.05 | x >= 0.95)) 
ppmcpairs <- cbind(ppmcpairs, misfit)
row.names(ppmcpairs)  <- c("Yens Q3", "OR", "OR-Diff", 
                          "RESID", "REDIFF")
pairs.index <- which(lower.tri(diag(I)), arr.ind = TRUE)
pairs.index <- apply(pairs.index, 1, function(x) paste(rev(x), collapse = ","))
colnames(ppmcpairs) <- c(paste0("Items(", pairs.index, ")"), "Misfit")
kbl(t(ppmcpairs), align = "c", booktabs = TRUE, caption = "PPMC Methods at the Item-Pairs Level", escape = FALSE)
```


```{r}
out_table <- rbind(out_table,
                   data.frame(seed, 
                        model = "HalfDrift",
                        diag.table, 
                        ACF = t(ppmc01),
                        PACF = ppmc02,
                        LPACF = ppmc03,
                        MSSD = ppmc04,
                        itcor = t(ppmc05),
                        itcor2 = t(ppmc06),
                        itcor3 = t(ppmc07),
                        q1 = t(ppmc08),
                        q1alt = t(ppmc09),
                        lpacf = t(ppmc10),
                        t(ppmc11),
                        t(ppmc12),
                        t(ppmc13),
                        t(ppmc14),
                        t(ppmc15)))
```

```{r export}
out_file <- paste0("Results_part2_", I, "_items.dat")

if (!file.exists(out_file)) {
  write.table(out_table, out_file, row.names = FALSE, col.names = TRUE)
} else {
  write.table(out_table, out_file, row.names = FALSE, col.names = FALSE, append = TRUE)
}
```

```{r}
rm(diag.table, fit, ppmcitem, ppmcpairs, ppmctest, responses, stan.diag,
   standata, maxRhat, misfit, nbfmi, nbulk, ndiv, nRhat, ntail, ntree,
   pairs.index, ppmc01, ppmc02, ppmc03, ppmc04, ppmc05, ppmc06, ppmc07, 
   ppmc08, ppmc09, ppmc10, ppmc11, ppmc12, ppmc13, ppmc14, ppmc15, run.time)
```

\clearpage

```{r clean}
rm(list = setdiff(ls(), c(lsf.str(), "i", "j", "seeds", "Nitems", "params")))
```

# References



