---
title: "Computing PPMC Methods of the Empirical Example: Self-Esteem"
author: "Sebastian Castro-Alvarez"
date: "`r format(Sys.Date(), '%B %d del %Y')`"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{amsmath}
output: 
  bookdown::pdf_document2:
    toc: false
    keep_tex: false
    number_section: false
bibliography: references.bib
csl: apa7.csl
link-citations: true
always_allow_html: true
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE, fig.height = 4, fig.width = 6, fig.pos = "!H",  
                      warning=FALSE, message=FALSE)

#rmarkdown::render("Rmarkdown/PGdata_PPMC.Rmd")
```

```{r pre-env}
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores() - 1)
# options(mc.cores = 2)
library(bayesplot)
color_scheme_set("darkgray")

source("../R/IRT_models.R")
source("../R/IRT_plots.R")
source("../R/PPMC.R")
source("../R/genTVDPCM.R")
source("../R/tvdpcm2stan.R")

set.seed(2022)
```

# The Empirical Data

In this document, we apply the posterior predictive model checking methods that we have developed for the TV-DPCM after fitting the TV-DPCM (and also the PCM) to an empirical data set. These data were retrieved from @Kossakowski2017 and was previously analyzed by @Wichers2016. The data come from a 57 years old male (at the time) that had been diagnosed with major depressive disorder. The participant completed up to 10 semi-random assessments per day for 239 days. During this period, the participant also followed a blind gradual reduction of their anti-depressant medication dosage. 

During the study, the participant completed several items that aimed to measure their emotional states. In total there were 12 items that were measured on a 7-point Likert scale. @Wichers2016 found 3 components after analyzing the data with PCA. The extracted components were positive affect, negative affect, and Self-Esteem. Previously, we fitted the TV-DPCM to each of these sets of items as well as a combination of the items of negative affect with Self-Esteem. In these analyses, we merged some response categories, given that, for some items, some response categories were never chosen. In general, after merging, the items ended up with 5 response options. However, for positive and negative affect, we also fitted the model one additional time where we merged the response categories into 3 response options. Additionally, we also fitted the PCM to each of the three scales after reducing the number of response options to 5. In what follows, we present display the PPMC methods that were applied to the fitted models.

```{r data}
PG_data <- read.csv("../ESMdata/ESMdata.csv")
PG_data <- PG_data[PG_data$resp_abort == 0, ]
PG_data <- PG_data[!is.na(PG_data$mood_down), ]

# Select variables
PG_data <- PG_data[, c("date", "phase", "concentrat", "beepno", "resptime_s", 
                       grep("mood_", names(PG_data), value = TRUE), 
                       grep("pat_", names(PG_data), value = TRUE),
                       grep("se_", names(PG_data), value = TRUE))]

# Transform Dates
PG_data$date       <- as.Date(PG_data$date, format = "%d/%m/%y")
PG_data$resptime_s <- as.POSIXct(paste(PG_data$date, PG_data$resptime_s))

# Select phases 1 and 2
PG_dataComplete <- PG_data
PG_data <- PG_dataComplete[PG_dataComplete$phase >= 1 & 
                             PG_dataComplete$phase <= 2, ]
```

```{r components}
pa <- c("mood_relaxed", "mood_satisfi", "mood_enthus", 
        "mood_cheerf", "mood_strong")
na <- c("mood_down", "mood_lonely", "mood_anxious", 
        "mood_guilty")
mu <- c("mood_irritat", "pat_restl", "pat_agitate")
se <- c("se_selflike", "se_ashamed", "se_selfdoub", "se_handle")
```

```{r addNA}
# Add ID dummy and time as numeric
PG_data$id   <- 1
# Create and ordered variable of the day number
PG_data$dayno <- as.numeric(PG_data$date) - as.numeric(PG_data$date[1])
# Combine the day number and the beep number
PG_data$time  <- PG_data$beepno + 16 * PG_data$dayno
# Include NA
PG_data_na <- esmpack:::expand.esm(data = PG_data, 
                                   id   = "id", 
                                   time = "time",
                                   tinterval = 1, 
                                   include   = c(mu, na, pa, se))
```

## TV-DPCM of Self-Esteem (SE, 3 categories, 3 items)

Here, we fitted the TV-DPCM to the 3 items of Self-Esteem. Response categories 5 to 7 were merged into one.

```{r data2stan}
responses <- PG_data_na[, se[-4]]
responses$se_ashamed  <- 6 - responses$se_ashamed
responses$se_selfdoub <- 6 - responses$se_selfdoub
responses[responses < 3] <- 3
responses[responses > 5]  <- 5
responses <- responses - 2

I  <- ncol(responses)
nT <- nrow(responses)
K  <- max(responses, na.rm = TRUE)

standata <- tvdpcm2stan_data(resp = as.matrix(responses),
                             I    = I,
                             K    = K,
                             nT   = nT,
                             n_knots  = 8,
                             s_degree = 3)
```

```{r sumscores, fig.cap = "Sumscores of Self-Esteem"}
index <- which(!is.na(responses[, 1]))[seq(1, sum(!is.na(responses[, 1])), 20)]
plot(index, rowMeans(responses)[index], yaxt = 'n', las = 1, type = "l", 
     ylab = "Meanscores SE", xlab = "Beep")
```

```{r model-load}
rdsfile <- "../Fits/SelfEwithNA_3items_ph1-2"

fit      <- readRDS(paste0(rdsfile, ".rds"))
```

```{r diag}
stan.diag <- monitor(extract(fit,
                             pars = c("beta", "theta", "lambda", 
                                      "sigma2", "pvar", "attractor"),
                             permuted = FALSE, inc_warmup = FALSE), 
                                 warmup = 0, print = FALSE)

ndiv  <- get_num_divergent(fit)     # number of divergent transitions
nbfmi <- get_low_bfmi_chains(fit)   # number of chains with a low bayesian fraction missing information 
ntree <- get_num_max_treedepth(fit) # number of transitions that exceeded the maximum treedepth
nbulk <- sum(stan.diag$Bulk_ESS < 100 * dim(fit)[2]) # number of parameters with low bulk ESS
ntail <- sum(stan.diag$Tail_ESS < 100 * dim(fit)[2]) # number of parameters with low tail ESS
          
            #max Rhat
maxRhat <- round(max(rhat(fit, pars = c("beta", "theta", "lambda", 
                                        "sigma2", "pvar", "attractor"))), 4)
nRhat   <- sum(rhat(fit, pars = c("beta", "theta", "lambda",
                                  "sigma2", "pvar", "attractor")) > 1.05)
```

### Convergence Checks

```{r diagtable}
diag.table <- matrix(c(nRhat, ndiv, length(nbfmi), ntree, nbulk, ntail), nrow = 1)
row.names(diag.table)  <- "Fitted Model"
colnames(diag.table) <- c("Rhat>1.05", "N. Divergent", "N. Low BFMI", 
                           "Excedeed Treedepth", "Low Bulk ESS", "Low Tail ESS")
kbl(diag.table, align = "c", booktabs = TRUE, caption = "HMC Diagnostics", escape = FALSE)
```

\clearpage

### Time Series of the Sumscores

```{r sums-ts, fig.cap = "Observed vs. Simulated Sumscores", results = 'hide'}
ppmc.sumscore.ts(object = fit, data = standata)
```

\clearpage

### Autocorrelation: 1st, 2nd, and 3rd

```{r acf, results = 'hide', fig.cap = "1st, 2nd, and 3rd autocorrelation of the sumscores"}
par(mfrow = c(3, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.acf(object = fit, data = standata, lag.max = 3)
```

\clearpage

### Partial Autocorrelation of the Observed Sumscores: AR(1) Residuals

```{r racf, results = 'hide', fig.cap = "1st Partial Autocorrelation of the Sumscores"}
ppmc.racf(object = fit, data = standata)
```

\clearpage

### Partial Autocorrelation of the Observed Sumscores: TV-DPCM Residuals

```{r lpacf-sum, results = 'hide', fig.cap = "1st Partial Autocorrelation of the Sumscores"}
time0 <- proc.time()
ppmc.lpacf(object = fit, data = standata, quiet = TRUE, sumscores = TRUE)
proc.time() - time0
```

\clearpage

### Mean Squared Successive Differences

```{r mssd, results = 'hide', fig.cap = "Mean Squared Succesive Differences of the Sumscores"}
ppmc.mssd(object = fit, data = standata)
```

\clearpage

### Item Scores Time Series

```{r item-ts, results = 'hide', fig.height = 6, fig.cap = "Observed vs. Simulated Item Scores"}
par(mfrow = c(I, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.item.ts(object = fit, data = standata, quiet = TRUE)
```

\clearpage

### Item-Total Correlation: Version 1

```{r itcor, results = 'hide', fig.height = 6, fig.cap = "Item-Total Correlation"}
par(mfrow = c(I, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.itcor(object = fit, data = standata, quiet = TRUE, method = "pearson")
```

\clearpage

### Item-Total Correlation: Version 2

```{r itcor2, results = 'hide', fig.height = 6, fig.cap = "Item-Total Correlation with AR(1) Residuals for the Total"}
par(mfrow = c(I, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.itcor2(object = fit, data = standata, quiet = TRUE, method = "pearson")
```

\clearpage

### Item-Total Correlation: Version 3

```{r itcor3, results = 'hide', fig.height = 6, fig.cap = "Item-Total Correlation with AR(1) Residuals for Item and Total Scores"}
par(mfrow = c(I, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.itcor3(object = fit, data = standata, quiet = TRUE)
```

\clearpage

### Yen's Q1

```{r q1, fig.height = 6, results = 'hide', fig.cap = "Yen's Q1 of the items of SE"}
time0 <- proc.time()
par(mfrow = c(I, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.Q1(object = fit, data = standata, quiet = TRUE)
proc.time() - time0
```

\clearpage

### Yen's Q1 Modified

```{r q1-alt, fig.height = 6, results = 'hide', fig.cap = "Modified Yen's Q1 of the items of SE"}
time0 <- proc.time()
par(mfrow = c(I, 1), mar = c(2, 3, 2, 2) + 0.1)
ppmc.Q1.alt(object = fit, data = standata, quiet = TRUE)
proc.time() - time0
```

\clearpage

### Partial Autocorrelation of the Observed Item Scores: TV-DPCM Residuals

```{r lpacf, results = 'hide', fig.cap = "1st Partial Autocorrelation of the Item Scores", fig.height = 6}
time0 <- proc.time()
par(mfrow = c(I, 1), mar = c(3, 4, 2, 2) + 0.1)
ppmc.lpacf(object = fit, data = standata, quiet = TRUE)
proc.time() - time0
```

\clearpage

### Yen's Q3

```{r q3, results = 'hide', fig.cap = "Yen's Q3 for the items of SE"}
time0 <- proc.time()
ppmc.Q3(object = fit, data = standata)$ppp
proc.time() - time0
```

\clearpage

### OR

```{r or, results = 'hide', fig.cap = "Odds Ratio for the items of SE"}
time0 <- proc.time()
ppmc.OR(object = fit, data = standata)$ppp
proc.time() - time0
```

\clearpage

### OR Modified

```{r ORdiff, results = 'hide', fig.cap = "Odds Ratio Difference between Two Halves of the Observations"}
time0 <- proc.time()
ppmc.ORDiff(object = fit, data = standata)$ppp
proc.time() - time0
```

\clearpage

### RESID

```{r resid, results = 'hide', fig.cap = "Absolute Item Covariance Residual of the items of SE"}
time0 <- proc.time()
ppmc.cov.resid(object = fit, data = standata)$ppp
proc.time() - time0
```

\clearpage

### RESID Difference

```{r rediff, results = 'hide', fig.cap = "Absolute Item Covariance Residual Difference between Two Halves of the Observations"}
time0 <- proc.time()
ppmc.cov.rediff(object = fit, data = standata)$ppp
proc.time() - time0
```

\clearpage

# References



